\levelstay{Hamiltonian approach}

\leveldown{Kirchhoff}

Here we analyze coupled harmonic resonators using the Hamiltonian method.
The material here is adapted from \cite{Sank:balanced_coupling:2025}.
We consider the system shown in Fig.~\ref{fig:coupled_resonators}, where two resonator are coupled through a capacitor $C_g$ and a mutual inducance $L_g$.
Kirchhoff's laws for this circuit produce the following equations of motion:
\begin{align}
  I_{C_a} + I_{L_a} + I_g &= 0 \nonumber \\
  I_{C_b} + I_{L_b} - I_g &= 0 \nonumber \\
  L_a \dot I_{L_a} + L_g \dot I_{L_b} &= V_a \nonumber \\
  L_b \dot I_{L_b} + L_g \dot I_{L_a} &= V_b
  \label{eq:coupled_oscillators_kirchhoff}
  \, .
\end{align}
In order to get useful equations of motion from Kirchhoff's laws, we need to work with all voltages or all currents.
We'll use voltages.
The capacitor currents $I_{C_a}$ and $I_{C_b}$ are easily related to voltage via the constitutive equation for a capacitor: $C_i \dot V_i = I_i$.
For example, the contitutive relation for the coupling capacitor gives $I_g = C_g (\dot V_a - \dot V_b)$.
Relating the inductor currents to voltages requires more work.
The bottom two of Eqs.~(\ref{eq:coupled_oscillators_kirchhoff}) can be written in matrix form as
\begin{equation*}
  \begin{pmatrix} V_a \\ V_b \end{pmatrix}
  = \underbrace{ \begin{pmatrix}
    L_a & L_g \\
    L_g & L_b
  \end{pmatrix}}_{T_L}
  \begin{pmatrix} \dot I_a \\ \dot I_b \end{pmatrix}
  \, .
\end{equation*}
Because we want to substitute voltage in where we have currents, we have to invert $T_L$,
\begin{align*}
  T_L^{-1}
  &= \frac{1}{L_a L_b - L_g^2} \begin{pmatrix}
    L_b & -L_g \\ -L_g & L_a
  \end{pmatrix} \\
  & \equiv \begin{pmatrix}
    1 / L_a' & -1 / L_g' \\ -1 / L_g' & 1 / L_b'
  \end{pmatrix}
\end{align*}
where
\begin{align*}
  L_a' & \equiv L_a - \frac{L_g^2}{L_b} \\
  \text{and} \quad
  L_b' & \equiv L_b - \frac{L_g^2}{L_a}
\end{align*}
are the inductances to ground for each resonator, including the inductance through the mutual.
Now we can express the inductor currents in terms of the voltages as
\begin{align*}
  \dot I_{L_a} &= \frac{V_a}{L_a'} - \frac{V_b}{L_g'} \\
  \dot I_{L_b} &= \frac{V_b}{L_b'} - \frac{V_a}{L_g'}
\end{align*}
allowing us to rewrite all the currents in the first two of Kirchhoff's laws in terms of $V_a$ and $V_b$:
\begin{align*}
  C_a \ddot V_a + \frac{V_a}{L_a'} - \frac{V_b}{L_g'} + C_g (\ddot V_a - \ddot V_b) &= 0 \\
  C_b \ddot V_b + \frac{V_b}{L_b'} - \frac{V_a}{L_g'} + C_g (\ddot V_b - \ddot V_a) &= 0 \, .
\end{align*}
Traditional analysis of circuits in the physics literature uses flux $\Phi$ and charge $Q$ instead of current and voltage, so defining $\Phi = \int V \, dt$, we can write our equations of motion as
\begin{align}
  \ddot \Phi_a (C_a + C_g) - C_g \ddot \Phi_b + \frac{\Phi_a}{L_a'} - \frac{\Phi_b}{L_g'} &= 0 \nonumber \\
  \ddot \Phi_b (C_b + C_g) - C_g \ddot \Phi_a + \frac{\Phi_b}{L_b'} - \frac{\Phi_a}{L_g'} &= 0
  \label{eq:coupled_equations_of_motion}
  \, .
\end{align}

\levelstay{Kirchhoff to Hamiltonian}

By inspection and a bit of fiddling around, one can check that equations (\ref{eq:coupled_equations_of_motion}) are generated by the Lagrangian
\begin{align}
  \mathcal{L}
  =& \underbrace{\frac{C_g}{2} \left(\dot \Phi_a - \dot \Phi_b \right)^2
   + \frac{C_a}{2} \dot \Phi_a^2 + \frac{C_b}{2} \dot \Phi_b^2}_\text{kinetic} \nonumber \\
  & \underbrace{- \frac{\Phi_a^2}{2 L_a'} - \frac{\Phi_b^2}{2 L_b'} + \frac{\Phi_a \Phi_b}{L_g'}}_\text{potential} \, .
  \label{eq:coupled_oscillators_lagrangian} % export
\end{align}
The canonical momenta conjugate to $\Phi_a$ and $\Phi_b$ are
\begin{align*}
  Q_a & \equiv \frac{\partial \mathcal{L}}{\partial \dot \Phi_a} = (C_a + C_g) \dot \Phi_a - C_g \dot \Phi_b \\
  Q_b & \equiv \frac{\partial \mathcal{L}}{\partial \dot \Phi_b} = (C_b + C_g) \dot \Phi_b - C_g \dot \Phi_a
  \, .
\end{align*}
or in matrix form
\begin{equation*}
  \begin{pmatrix} Q_a \\ Q_b \end{pmatrix}
  =
  \underbrace{
    \begin{pmatrix}
       (C_a + C_g) & -C_g \\
       -C_g & (C_b + C_g)
    \end{pmatrix}
  }_{T_C}
  \begin{pmatrix} \dot \Phi_a \\ \dot \Phi_b \end{pmatrix}
  \, .
\end{equation*}
The matrix $T_C$ is just the capacitance matrix of the circuit.
Its inverse is
\begin{align*}
  T_C^{-1} =& \frac{1}{C_a C_b + C_g (C_a + C_b)}
  \begin{pmatrix}
    (C_b + C_g) & C_g \\
    C_g & (C_a + C_g)
  \end{pmatrix} \\
  \equiv& \begin{pmatrix}
    1 / C_a' & 1 / C_g' \\
    1 / C_g' & 1 / C_b'
  \end{pmatrix}
\end{align*}
Note that $C_a'$ and $C_b'$ are simply the total capacitances to ground for each resonator.

The Hamiltonian function $H$ for the system is defined formally by the equation
\begin{equation*}
  H \equiv \left( \sum_{i = \{a, b\}} Q_i \dot \Phi_i \right) - \mathcal{L}
\end{equation*}
where $\Phi_i$ are the coordinates and $Q_i$ are the conjugate momenta, but where we have to replace the $\dot \Phi$'s in both the $Q \dot \Phi$ terms and in $\mathcal{L}$ with $\Phi$'s and $Q$'s.
To do this, first note that the kinetic term in the Lagrangian can be expressed as\footnote{Because matrix transposition and inversion commute, and because the matrix $T_C$ is symmetric, we can bring $T_C^{-1}$ from the bra onto the ket for free.}
\begin{align*}
  \mathcal{L}_\text{kinetic}
  =& \frac{1}{2} \bbraket{\dot \Phi}{T_C}{\dot \Phi} \\
  =& \frac{1}{2} \bbraket{T_C^{-1} Q}{T_C}{T_C^{-1} Q} \\
  =& \frac{1}{2} \bbraket{Q}{T_C^{-1}}{Q}
  \, .
\end{align*}
Note also that $\sum_i \dot \Phi_i Q_i = \bbraket{Q}{T_C^{-1}}{Q}$, so we can write the Hamiltonian as
\begin{align}
  H
  =& \left( \sum_{i = \{a, b\}} Q_i \dot \Phi_i \right) - \mathcal{L} \nonumber \\
  =& \bbraket{Q}{T_C^{-1}}{Q} - \left( \mathcal{L}_\text{kinetic} + \mathcal{L}_\text{potential} \right) \nonumber \\
  =& \bbraket{Q}{T_C^{-1}}{Q} - \left( \frac{1}{2} \bbraket{Q}{T_C^{-1}}{Q} + \mathcal{L}_\text{potential} \right) \nonumber \\
  =& \frac{Q_a^2}{2 C_a'} + \frac{Q_b^2}{2 C_b'} + \frac{\Phi_a^2}{2 L_a'} + \frac{\Phi_b^2}{2 L_b'} \nonumber \\
  &+ \underbrace{\frac{Q_a Q_b}{C_g'} - \frac{\Phi_a \Phi_b}{L_g'}}_{\text{coupling Hamiltonian }H_g}
  \, .
\end{align}

\levelstay{Dimensionless variables}

We will now simplify this Hamiltonian so that we can easily find its normal modes.
First, we define
\begin{align*}
  X_i &\equiv \frac{1}{\sqrt{2 \hbar}} \frac{1}{\sqrt{Z_i'}} \Phi_i \\
  Y_i &\equiv \frac{1}{\sqrt{2 \hbar}} \sqrt{Z_i'} Q_i
\end{align*}
where $Z_i' \equiv \sqrt{L_i' / C_i'}$, and we've added the constant $\hbar$ with dimensions of action to make $X$ and $Y$ dimensionless.
This entire analysis has been classical, and in the classical case $\hbar$ can be thought of as \emph{anything} with dimensions of action.
Of course, in the quantum case, we should simply think of $\Phi$, $Q$, $X$, and $Y$ as operators and $\hbar$ as Planck's constant.

In the new coordinates, the Hamiltonian is
\begin{align}
  H / \hbar
  =& \omega_a' \left(X_a^2 + Y_a^2 \right)
  +  \omega_b' \left(X_b^2 + Y_b^2 \right) \nonumber \\
  &+ 2 \frac{1}{C_g' \sqrt{Z_a' Z_b'}} Y_a Y_b
   - 2 \frac{\sqrt{Z_a Z_b}}{L_g'} X_a X_b
\end{align}
where $\omega_i' \equiv 1 / \sqrt{L_i' C_i'}$ are called \textbf{partial frequencies} and play an important role in the analysis of the system, particularly when making approximations.

\levelstay{Rotating modes}

Finally we define
\begin{align}
  a &\equiv X_a + i Y_a \nonumber \\
  b &\equiv X_b + i Y_b
\end{align}
to arrive at
\begin{align}
  H / \hbar
  &= \omega_a' a^* a + \omega_b' b^* b \nonumber \\
  & - \frac{1}{2} \frac{1}{C_g' \sqrt{Z_a' Z_b'}} (ab + a^* b^* - a b^* - a^* b) \nonumber \\
  &- \frac{1}{2} \frac{\sqrt{Z_a' Z_b'}}{L_g'} (ab + a^* b^* + a^* b + a b^*)
  \, .
\end{align}
The stars indictate Hermitian conjugation, which in the classical case reduces to complex conjugation.
The coupling term can be reorganized into the form:
\begin{align}
  H_g / \hbar =
  &- \left( a b + a^* b^* \right)
    \frac{1}{2} \left(
      \frac{1}{C_g' \sqrt{Z_a' Z_b'}} + \frac{\sqrt{Z_a' Z_b'}}{L_g'}
    \right) \nonumber \\
  &+ \left( a b^* + a^* b \right)
    \frac{1}{2} \left(
      \frac{1}{C_g' \sqrt{Z_a' Z_b'}} - \frac{\sqrt{Z_a' Z_b'}}{L_g'}
    \right) \nonumber \\
  &= -g_+ (ab + a^* b^*) + g_- (ab^* + a^* b)
\end{align}
where we defined
\begin{align}
  g_c \equiv \frac{1}{2} \frac{1}{C_g' \sqrt{Z_a' Z_b'}} &\qquad
  g_l \equiv \frac{1}{2} \frac{\sqrt{Z_a' Z_b'}}{L_g'} \nonumber \\
  \text{and} \qquad \qquad
  g_+ \equiv g_c + g_l &\qquad g_- \equiv g_c - g_l
  \, ,
\end{align}
which is the starting point in the main text.

\levelstay{Discussion}

Evidently, there are two \emph{aspects} to the coupling.
The $g_-$ term conserves excitation number, while the $g_+$ term does not and is dropped under the RWA.
Therefore, if we engineer the coupling such that $g_+$ is identically zero, then the RWA will be exact, i.e. the coupling Hamiltonian in the rotating frame becomes
\begin{equation}
  \bar{H}_g / \hbar = g_- \left( \bar{a} \bar{b}^* e^{-i (\omega_a' - \omega_b')}+ \text{h.c.} \right)
  \, .
\end{equation}
This observation, that a purely number conserving interaction can be exactly realized in electromagnetic circuits, is the second main result of the paper.

The condition $g_+ = 0$ means $g_c = -g_l$, i.e. the capacitive and inductive coupling must be of equal magnitude but opposite sign, which we call ``balanced coupling''.
Balanced coupling can also be expressed as the condition $\sqrt{-L'_g / C'_g} = \sqrt{Z_a' Z_b'}$, i.e. the coupling impedance is equal to the geometric mean of the impedances of the two resonators.
Recall that a negative value for $L_g$ means that the mutual inductance must be arranged so that a positive value of $\dot{I}_{L_a}$ produces a negative contribution to $V_b$.
If the capacitive coupling is positive, then balanced coupling requires the mutual inductance to be negative.
It should be noted that with balanced coupling, in order to achieve a certain total coupling strength $g$, both $g_c$ and $g_l$ would be \emph{half} of what would be needed in the case of e.g. purely capacitive coupling.
Also note that strong inductive coupling has been shown even with low impedance qubits \cite{Neill:gmon:2014}.

\levelstay{Equations of motion}\label{sec:equations_of_motion}

We take a brief aside to comment on the equations of motion arising from Eq.~(\ref{eq:ham_coupled_res}), which can be expressed in matrix form as
\begin{equation}
  \frac{d}{dt}
  \left( \begin{array}{c} a \\ b \\ a^* \\ b^* \end{array} \right)
  = -i \underbrace{\left( \begin{array}{cccc}
    \omega_a' & g_- & 0 & -g_+ \\
    g_- & \omega_b' & -g_+ & 0 \\
    0 & g_+ & -\omega_a' & -g _- \\
    g_+ & 0 & -g_- & -\omega_b'
  \end{array} \right)}_M
  \left( \begin{array}{c} a \\ b \\ a^* \\ b^* \end{array} \right) \, .
\end{equation}
The coupling $g_-$ connects $a$ to $b$ and $a^*$ to $b^*$, i.e. it acts within the upper and lower diagonal 2x2 sub-blocks of $M$.
On the other hand, the coupling $g_+$ couples $a$ to $b^*$ and $b$ to $a^*$ as it lies on the anti-diagonal of $M$.
The structure of $M$ and the roles of the two aspects of coupling can be clearly seen by writing $M$ in algebraic form as
\begin{equation}
  M = -i \left[
    \sigma_z \otimes
      \left(
        g_- \sigma_x + \frac{\Delta}{2} \sigma_z + \frac{S}{2} \mathbb{I}
      \right)
    -i g_+ (\sigma_y \otimes \sigma_x)
  \right]
\end{equation}
where $S \equiv \omega_a' + \omega_b'$ and $\Delta \equiv \omega_a' - \omega_b'$.

Dropping the $g_+$ terms in the Hamiltonian is equivalent to dropping the antidiagonal of $M$.
Appendix \ref{appendix:eigenvalues} compares the full eigenvalues of $M$ with the eigenvalues of $M$ under the RWA.


\levelstay{Old stuff}

\begin{align}
  H
  &= \frac{\Phi_1^2}{2 L_1} + \frac{\Phi_2^2}{2 L_2} \\
  &+ \frac{Q_1^2}{2 C_1} + \frac{Q_2^2}{2 C_2} \\
  &+ \frac{Q_1 Q_2}{C_g''}
\end{align}
where
\begin{align}
  C_g'' &\equiv \frac{C_1' C_2' - C_g^2}{C_g} \\
  C_i' &\equiv C_i + C_g \, .
\end{align}
We transform the Hamiltonian in two steps.
First we define
\begin{align*}
  X_i &\equiv \frac{1}{\sqrt{2 \hbar}} \frac{1}{\sqrt{Z_i}} \Phi_i \\
  Y_i &\equiv \frac{1}{\sqrt{2 \hbar}} \sqrt{Z_i} Q_i
\end{align*}
where $Z_i \equiv \sqrt{L_i / C_i''}$.
With these new variables, the Hamiltonian is
\begin{equation}
  H/\hbar =
    \omega_1 \left(X_1^2 + Y_1^2) \right)
  + \omega_2 \left(X_2^2 + Y_2^2) \right)
  + 2 \left( \frac{C_1'' C_2''}{C_g''} \right) \sqrt{\omega_1'' \omega_2''} Y_1 Y_2 \, .
\end{equation}
Note that $[X, Y] = i/2$ so Hamilton's equations of motion are a little different than usual.
Anyway, we next use $a = X_1 + i Y_1$ and $b = X_2 + i Y_2$ to get
\begin{equation}
  H / \hbar =
    \omega_1'' a^\dagger a
  + \omega_2'' b^\dagger b
  - g (a - a^\dagger) (b - b^\dagger)
\end{equation}
where
\begin{equation}
  g \equiv \frac{1}{2} \frac{\sqrt{C_1'' C_2''}}{C_g''} \sqrt{\omega_1'' \omega_2''} \, .
\end{equation}
Hamilton's equations of motion in this representation are
\begin{equation*}
  \dot{a} = -i \frac{\partial H/\hbar}{\partial a^\dagger}
  \quad \text{and} \quad
  \dot{a}^\dagger = i \frac{\partial H/\hbar}{\partial a}
\end{equation*}
and similarly for $b$.
Therefore the equations of motion for the system are
\begin{align*}
  \dot a &= -i \omega_1'' a -i g b + i g b^\dagger \\
  \dot b &= -i \omega_2'' b -i g a + i g a^\dagger \, .
\end{align*}
These equations can be written in matrix form as
\begin{equation}
  \frac{d}{dt}
  \left( \begin{array}{c}
    a \\ b \\ a^\dagger \\ b^\dagger
  \end{array} \right)
  =
  -i \left( \begin{array}{cccc}
    \omega_1'' & g & 0 & -g \\
    g & \omega_2'' & -g & 0 \\
    0 & g & - \omega_1'' & -g \\
    g & 0 & -g & - \omega_2''
  \end{array} \right)
  \left( \begin{array}{c}
    a \\ b \\ a^\dagger \\ b^\dagger
  \end{array} \right)
  \, .
\end{equation}
The matrix can be conveniently expressed as
\begin{equation}
  \sigma_z \otimes \left(
    g \sigma_x + \frac{\Delta}{2} \sigma_z + \frac{S}{2} \mathbb{I}
  \right)
  -i g \left( \sigma_y \otimes \sigma_x \right)
\end{equation}
where $\Delta \equiv \omega_1'' - \omega_2''$ and $S \equiv \omega_1'' + \omega_2''$.
The eigenvalues are
\begin{equation}
  \lambda^2 =
  \pm \frac{1}{2}
  \left(
    \omega_1''^2 + \omega_2''^2
    \pm \sqrt{(\omega_1''^2 - \omega_2''^2)^2 + 16 g^2 \omega_1'' \omega_2''}
  \right)
\end{equation}

In the rotating wave approximation, we drop the $ab$ and $a^\dagger b^\dagger$ terms in the Hamilonian.
Dropping those terms is equivalent to dropping the $-i g \left( \sigma_y \otimes \sigma_x \right)$ term in the algebraic representation!
Doing that and computing the eigenvalues again gives
\begin{equation}
  \lambda_\pm =
  \frac{1}{2} \left( S \pm \sqrt{\Delta^2 + 4 g^2} \right) \, .
\end{equation}
