\documentclass{article}

\input{../TeX/packages.tex}
\input{../TeX/macros.tex}

\title{Harmonic Oscillator}
\author{Daniel Sank \\ \small{University of California Santa Barbara} \\ \small{Presently Google Quantum AI}}
\date{23 February 2009}

\begin{document}
\maketitle

\section{Impulse Response}

The equation for a damped harmonic oscillator is \begin{equation}
\ddot{\phi}(t)+2\beta\dot{\phi}(t)+\omega_{0}^{2}\phi(t)=j(t) \end{equation}
In the usual way we turn this into an abstract equation \begin{equation}
(D_{t}^{2}+2\beta D_{t}+\omega_{0}^{2})\ket{\phi}=\ket{j} \end{equation}
Our Fourier convention will be
\begin{equation}
\braket{t}{\omega}=e^{i\omega t}
\end{equation}
which is the common choice in electrical engineering.
However, it is the \emph{opposite} of what is normally used in physics, notably in Schrodinger's equation.
With this convention $-iD_t$ is Hermitian.
Going to the $\ket{\omega}$ basis we get
\begin{align}
\bbraket{\omega}{D_t^2 + 2\beta D_t + \omega_0^2}{\phi} &= \braket{\omega}{j} \\
(-\omega^{2} + 2 i \beta\omega + \omega_{0}^{2}) \phi(\omega) & = j(\omega)\\
\phi(\omega) & = \frac{j(\omega)}{-\omega^{2} + 2 i \beta \omega + \omega_{0}^{2}} \, .
\end{align}
Taking $j(t)=A\delta(t-t_{0})$ we get $\braket{\omega}{j} = A e^{-i \omega t_{0}}$.
We can factor the denominator as $-(\omega-\omega_{+})(\omega-\omega_{-})$
where
\begin{equation}
\omega_{+} = i \beta + \omega_0' \qquad \omega_{-} = i \beta - \omega_0' \label{eq:omega_plus_minus}
\end{equation}
and
\begin{equation}
\omega_0' = \omega_0 \sqrt{ 1 - \left( \beta / \omega_0 \right) ^2 } \, .
\end{equation}
Therefore we have \begin{equation}
\phi(\omega) = \frac{-A e^{-i \omega t_0}}{(\omega - \omega_{+})(\omega-\omega_{-})} \, .
\end{equation}
Upon Fourier transform
\begin{equation}
\phi(t) = 
\int\frac{-Ae^{i \omega(t - t_0)}}{(\omega-\omega_{+})(\omega-\omega_{-})} \,
\frac{d\omega}{2\pi} \, .
\end{equation}
Consider the case $t < t_0$.
In that case we close the contour in the lower half plane.
Since both poles $\omega_\pm$ are in the upper half plane the integral is zero.
This reflects causality: the oscillator can't know about the impulse before it happens.
For the case $t > t_{0}$ we close in the upper half plane.
Now the contour encloses the poles and we will get a nonzero result.
Let $\tau \equiv t - t_0$.
Then the residue theorem gives
\begin{align}
\phi(t) &=
\frac{-A 2\pi i}{2 \pi}
\left(
\frac{e^{i \omega_+ \tau}}{\omega_+ - \omega_-} +
\frac{e^{i \omega_- \tau}}{\omega_- - \omega_+}
\right) \\
&= \frac{A}{\omega_0'} e^{- \beta \tau} \sin(\omega_0' \tau)
\end{align}
We see that the system oscillates with a frequency close to $\omega_0$ and decays with a rate given by the amplitude damping parameter $\beta$.
Note that this solution is valid only when $\omega_0 > \beta$.

In the limit $\beta \ll \omega_0$ we find $E(t) = E_{0}e^{-2 \beta t}$, where $E_0=\frac{1}{2} m A^{2}$ is the initial energy of the system.
Note that $A$ is the initial velocity.

Defining the quality factor $Q$ by
\begin{equation}
Q \equiv \frac{\textrm{Energy stored}}{\textrm{Energy dissipated per radian}} \, ,
\end{equation}
we find
\begin{equation}
Q
= \frac{E(t)}{dE/d\textrm{rad}}
= \frac{E(t)}{-(dE/dt)(dt/d\textrm{rad})}
= \frac{E(t)}{2\beta E(t)\frac{1}{\Delta\omega}}
= \frac{\Delta\omega}{2\beta}\approx\frac{\omega_{0}}{2\beta} \, .
\end{equation}
This relation is also written
\begin{equation}
Q \approx \omega_0 / \kappa
\end{equation}
where $\kappa$ is the energy decay rate of the oscillator.


\section{Frequency Response}

Now we investigate the case in which $j(t)=A\cos(\Omega t)$.
In this case
\begin{equation}
j(\omega)=\frac{A}{2}(2\pi)\left[\delta(\omega-\Omega)+\delta(\omega+\Omega)\right] \, .
\end{equation}
Therefore, by modifying equations from the previous section we have
\begin{align}
\phi(\omega)
& = \frac{-A}{2}(2\pi) \int \frac{d\omega}{2\pi} e^{i \omega t} \frac{\delta(\omega-\Omega)+\delta(\omega+\Omega)}{\omega^{2} - 2i\beta\omega - \omega_0^2} \\
& =
\frac{-A}{2}
\left[
\frac{e^{i \Omega t}}{\Omega^{2} - 2i\beta\Omega - \omega_0^2} +
\frac{e^{-i \Omega t}}{\Omega^2 + 2i\beta\Omega - \omega_0^2}
\right] \\
&= \Re \left[ e^{i \Omega t} \frac{-A}{\Omega^2 - 2i\beta\Omega - \omega_0^2}
\right] \label{eq:phasor_form}
\end{align}
This is a very special form known as {}``phasor'' form.
The idea of phasors is that in a linear system driven by a sinusoidal drive, all variables will oscillate with the same frequency as the drive.
Any dynamical variable, in our case $\phi$, differs from the drive only in amplitude and phase.
We see this here: the time dependence of $\phi$ is sinusoidal at angular frequency $\Omega$.
The crucial relation is that if we have two signals at the same frequency which can be expressed as \begin{equation}
f(t) = \Re[ e^{i\omega t} \hat{f}] \qquad
g(t) = \Re[ e^{i\omega t} \hat{g}]
\end{equation}
where $\hat{f}$ and $\hat{g}$ are complex numbers, then we have \begin{equation}
\langle f(t)g(t)\rangle_{t}=\frac{1}{2}\Re[\hat{f}^{*}\hat{g}] \end{equation}
where the left hand is a time average.
Our expression for $\phi(t)$ has this form if we take
\begin{equation}
\hat{\phi} = \frac{-A}{\Omega^{2} - 2i\beta\Omega - \omega_0^2} \, .
\end{equation}
Therefore, \begin{equation}
  \langle\phi(t)^{2}\rangle_{t}
  = \frac{1}{2} \frac{A^{2}}{(\Omega^{2}-\omega_0^{2})^{2}+(2\beta\Omega)^2}
  \, .
\end{equation}
This function maximizes for $\Omega=\pm\sqrt{\omega_{0}^{2}-2\beta^{2}}$.
Note that this is not the same as the free oscillation frequency found from the impulse response.

\subsection{Resonance}

We define the resonance frequency as that frequency at which power flow from the drive to the system is unidirectional.
This happens when $\dot{\phi}(t)$ is in phase with $j(t)$.
Velocity and position are a quarter cycle out of phase, so the resonance occurs when the position is a quarter cycle shifted from the drive.
In other words, the resonance happens when the denominator of the phasor is imaginary, which occurs for
\begin{equation}
(\text{resonance}) \qquad \Omega = \pm \omega_0 \, .
\end{equation}

\subsection{Phase shift}

It is useful to look at the relative phase shift between the drive and the response.
To do this we need to find the argument of $\hat{\phi}$.
Note that
\begin{equation}
f(t)
= \Re[e^{i \Omega t}\hat{f}]
= \Re[e^{i \Omega t}|f|e^{i\theta}]
=|f|\cos(\Omega t + \theta)]
\end{equation}
where $\theta$ is the phase of $\hat{f}$.
Therefore, a positive $\theta$ indicates that the response leads the source, and a negative $\theta$ indicates that the response lags the source.

\subsection{Lorentzian approximation}

We now express the resonance in the so-called Lorentzian form.
\begin{align}
\phi(t)
&= \Re \left[ e^{i\Omega t} \frac{-A}{\Omega^2 - \omega_0^2 - 2i\beta \Omega} \right] \nonumber \\
&= \frac{A \cos \left( \Omega t + \delta \right)}{\sqrt{\left( \Omega^2 - \omega_0^2 \right)^2 + (2 \beta \Omega)^2}}
\end{align}
where $\delta$ is a phase we don't care about.
We can re-express this approximately as
\begin{equation}
\phi(t) \approx \frac{(A \beta / \Omega) \cos(\Omega t + \delta)}{\left( \Omega - \omega_0 \right)^2 + 2\beta^2} \, .
\end{equation}


\section{Phasor form}

It is possible to find approximation equations of motion for the driven oscillator in the case that the spectral content of the drive is reasonably well localized around the resonance frequency.
Consider the frequency domain equation of motion
\begin{equation}
  \phi(\omega) = \frac{-J(\omega)}{\omega^2 - 2 i \beta \omega - \omega_0^2}
  = - \frac{J(\omega)}{(\omega - \omega_+)(\omega - \omega_-)} \, .
\end{equation}
Using the method of partial fractions, we can rewrite the double pole as a sum of single poles
\begin{equation}
  \frac{1}{(\omega - \omega_+)(\omega - \omega_-)}
  = \frac{X}{\omega - \omega_+} + \frac{Y}{\omega - \omega_-}
\end{equation}
with unknown constants $X$ and $Y$.
Multiplying through by $(\omega - \omega_+)(\omega - \omega_-)$ and solving gives
\begin{align*}
  X =& \frac{1}{\omega_+ - \omega_-} =   \frac{1}{2 \omega_0'} \\
  Y =& \frac{1}{\omega_- - \omega_+} = - \frac{1}{2 \omega_0'}
\end{align*}
so
\begin{equation}
  \phi(\omega) = - \frac{J(\omega)}{2 \omega_0'}
  \left[ \frac{1}{\omega - \omega_+} - \frac{1}{\omega - \omega_-} \right] \, .
\end{equation}
If the drive is a narrow band signal near a central frequency $\Omega$, we can write it as
\begin{equation}
  J(t) = I(t) \cos(\Omega t) - i Q(t) \sin(\Omega t) \, ,
\end{equation}
in which case we have
\begin{align*}
  \tilde{J}(\omega)
  &= \frac{1}{2} \left(
      \tilde{I}(\omega - \Omega) + i \tilde{Q}(\omega - \Omega)
    + \tilde{I}(\omega + \Omega) - i \tilde{Q}(\omega + \Omega)
  \right) \\
  &= \frac{1}{2} \left(
      \tilde{z}(\omega - \Omega) + \tilde{z}(-\omega - \Omega)^*
  \right)
\end{align*}
where $z(t) \equiv I(t) - i Q(t)$ and we've used the fact that $\tilde{I}(\omega)^* = \tilde{I}(-\omega)$ and $\tilde{Q}(\omega)^* = \tilde{Q}(-\omega)$ because $I(t)$ and $Q(t)$ are real.
Combining results, we have
\begin{align*}
  \tilde{\phi}(\omega)
  =& - \frac{1}{4 \omega_0'}
  \left( \tilde{z}(\omega - \Omega) + \tilde{z}(-\omega - \Omega)^* \right)
  \left(
    \frac{1}{\omega - \omega_+} - \frac{1}{\omega - \omega_-}
  \right) \\
  =& - \frac{1}{4 \omega_0'}
  \left(
      \frac{\tilde{z}(\omega - \Omega)}{\omega - \omega_+}
    - \frac{\tilde{z}(-\omega - \Omega)^*}{\omega - \omega_-}
    \underbrace{
      - \frac{\tilde{z}(\omega - \Omega)}{\omega - \omega_-}
      + \frac{\tilde{z}(-\omega - \Omega)^*}{\omega - \omega_+}
    }_\text{small}
  \right)
\end{align*}
The second two terms in parentheses are small because e.g. $\tilde{z}(\omega - \Omega)$ is large when $\omega$ is near $+\Omega$, but $1 / (\omega - \omega_-)$ is large for $\omega$ near $-\Omega$.
Keeping only the larger terms, we get
\begin{equation}
  \tilde{\phi}(\omega) =
  - \frac{1}{4 \omega_0'}
  \left(
      \frac{\tilde{z}(\omega - \Omega)}{\omega - \omega_+}
    - \frac{\tilde{z}(-\omega - \Omega)^*}{\omega - \omega_-}
  \right) \, .
\end{equation}
Note that $\tilde{\phi}(-\omega) = \tilde{\phi}(\omega)^*$ as appropriate for $\phi(t) \in \mathbb{R}$.
Because of that symmetry, we can write it as a Fourier transform over only positive frequencies:
\begin{align*}
  \phi(t)
  &= 2 \Re \int_0^\infty \frac{d\omega}{2\pi} \, e^{i \omega t} \tilde{\phi}(\omega) \\
  &= 2 \Re \int_0^\infty \frac{d\omega}{2\pi} \, e^{i \omega t} \left( \frac{-1}{4 \omega_0'} \right)
    \left(
      \frac{\tilde{z}(\omega - \Omega)}{\omega - \omega_+}
      - \underbrace{\frac{\tilde{z}(-\omega - \Omega)^*}{\omega - \omega_-}}_\text{small}
    \right) \\
  (\text{define } \omega \equiv \Omega + \delta \omega)
  &= \Re
  \left(
    e^{i \Omega t}
    \int_{-\Omega}^\infty \frac{d \delta \omega}{2\pi} \,
    e^{i \delta \omega t}
    \left( \frac{-1}{2 \omega_0'} \right)
    \frac{\tilde{z}(\delta \omega)}{\delta \omega - \Delta - i \beta}
  \right)
\end{align*}
where $\Delta \equiv \omega_0' - \Omega$.
If $\tilde{z}$ is band limited to $[-B, B]$ where $B \ll \omega_0'$, then we can rewrite the limits of the integration as $\pm \infty$:
\begin{equation}
  \phi(t) = \Re
  \left(
    e^{i \Omega t}
    \underbrace{
      \int_{-\infty}^\infty \frac{d \delta \omega}{2\pi} \,
      e^{i \delta \omega t}
      \left( \frac{-1}{2 \omega_0'} \right)
      \frac{\tilde{z}(\delta \omega)}{\delta \omega - \Delta - i \beta}
    }_{\psi(t)}
  \right)
\end{equation}
The underbraced part $\psi(t)$ is the part of the oscillator's motion that comes on top of a free oscillation at frequency $\Omega$.
In other words, we're in the rotating frame of the drive.
By inspection, the time dynamics of $\psi(t)$ are governed by the equation
\begin{equation}
  \dot{\psi}(t) = \left(\frac{-i}{2 \omega_0'}\right) z(t) + (i \Delta - \beta ) \psi(t)
\end{equation}

\section{Rotating Mode Formalism}

There's an interesting way to write down the equations for a harmonic oscillator that's useful when you have coupled problems or problems with time varied parameters.
An as example consider a parallel LRC circuit.
As dynamical variables take the flux in the inductor, $\Phi$, and the charge on the capacitor, $Q$.
See figure ??.
The two first order equations of motion are\begin{equation}
\dot{\Phi}=Q/C\qquad\dot{Q}+Q/(RC)=-\Phi/L \end{equation}
Note that if $R\rightarrow\infty$ these can be written in a nearly Hamiltonian form \begin{equation}
\frac{d}{dt}\left[ \begin{array}{c} \Phi \\ Q\end{array} \right] =
\left[\begin{array}{cc} 0 & 1/C \\ -1/L & 0 \end{array} \right]
\left[\begin{array}{c} \Phi\\ Q\end{array}\right] \end{equation}
If the off-diagonal elements were equal to one another this would be a true Hamiltonian.
This is easy to do if we simply re-scale the variables, \begin{equation}
X \equiv \left( C/L \right)^{1/4} \Phi \qquad Y \equiv \left( L/C \right)^{1/4} Q \end{equation}
Now we find, in the case that $R\rightarrow\infty$ \begin{displaymath}
\frac{d}{dt}\left[ \begin{array}{c} X \\ Y \end{array} \right] = \omega_{0} \left[ \begin{array}{cc} 0 & 1 \\ -1 & 0 \end{array}\right]
\left[ \begin{array}{c} X \\ Y \end{array} \right] \end{displaymath}
where $\omega_{0}\equiv1/\sqrt{LC}$.
This is great because we have the anti-symmetric Hamiltonian matrix.
Adding the loss back in gives \begin{displaymath}
\frac{d}{dt} \left[ \begin{array}{c} X \\ Y \end{array} \right] =
\omega_0 \left[ \begin{array}{cc} 0 & 1 \\ -1 & -1/Q \end{array} \right]
\left[ \begin{array}{c} X \\ Y \end{array} \right] \end{displaymath}
where $Q\equiv \omega_0 R C$.

We can make one further simplification if we want to get rid of the $\omega_0$ out front.
Make one more rescaling,\begin{equation}
\xi\equiv\omega_{0}t \end{equation}
Then \begin{equation}
\frac{d}{dt}=\omega_{0}\frac{d}{d\xi}\end{equation}
Therefore the equations of motion for $X$ and $Y$ are \begin{equation}
\frac{d}{d\xi}\left[\begin{array}{c} X\\ Y\end{array}\right] = 
\left[\begin{array}{cc} 0 & 1\\ -1 & -1/Q \end{array}\right]
\left[\begin{array}{c} X\\ Y\end{array}\right]\end{equation}

\subsection{Normal Modes}
We can now find the normal modes of this equation.
To do so we must find the eigenvalues and eigenvectors of the matrix, \begin{equation}
\left[ \begin{array}{cc} 0 & 1 \\ -1 & -\epsilon \end{array} \right] \end{equation}
where here $\epsilon \equiv 1/Q$.
The eigenvalue equation is \begin{equation}
\lambda^2 + \epsilon \lambda + 1 = 0 \end{equation}
from which the eigenvalues are found to be \begin{equation}
\lambda_{\pm} = \pm i \sqrt{1 - \frac{\epsilon^2}{4}} - \frac{\epsilon}{2} \end{equation}
Now, we want to find eigenvectors $a^+$ and $a^-$ that satisfy the equations \begin{equation}
\dot{a}^+=\lambda_+ a^+ \quad \textrm{and} \quad \dot{a}^-=\lambda_- a^- \end{equation}
where the overdots denote differentiation with respect to $\xi$.
Write $a$ (denoting here either one of the eigenvectors) as $a = X+\alpha Y$.
Then taking the derivative with respect to $\xi$ of both sides gives \begin{eqnarray}
\dot{a} &=& \dot{X}+\alpha \dot{Y}\\
&=& Y+\alpha(-X-\epsilon Y)\\
&=& -\alpha \left( X+\left[ \epsilon - \frac{1}{\alpha} \right] Y \right) \end{eqnarray}
Since we're expecting $a$ to satisfy $\dot{a} = \lambda a$ we get \begin{eqnarray}
\dot{a} &=& \lambda a \\
-\alpha \left( X+\left[ \epsilon - \frac{1}{\alpha} \right] Y \right) &=& \lambda \left( X + \alpha Y \right) \\
\rightarrow \quad \lambda=-\alpha &\quad& 1-\epsilon \alpha = \lambda \alpha \end{eqnarray}
where the last line follows by equating coefficients of $X$ and $Y$.
We can check that these relations are self consistent by plugging the first into the second \begin{eqnarray}
1+\epsilon \lambda &=& -\lambda^2 \\
1+\epsilon \left( -\frac{\epsilon}{2} \pm i \sqrt{1-\frac{\epsilon^2}{4}} \right) &=&
-\left(-\frac{\epsilon}{2}\pm i \sqrt{1-\frac{\epsilon^2}{4}} \right) \\
1-\frac{\epsilon^2}{2}\pm i \epsilon \sqrt{\cdots} &=&
-\frac{\epsilon^2}{4}\pm i \epsilon\sqrt{\cdots} + 1 -\frac{\epsilon^2}{4}\\
0 &=& 0 \end{eqnarray}
so the relations are indeed consistent.

Now we can write down the eigenvectors along with their time dependence, \begin{eqnarray}
a^- = \frac{1}{\sqrt{2}}\left( X -\lambda_- Y \right) &\qquad&
a^+ = \frac{1}{\sqrt{2}}\left( X -\lambda_+ Y \right) \\
a^-(t) = a^-(0) e^{\omega_0 \lambda_- t} &\qquad& a^+(t) = a^+(0) e^{\omega_0 \lambda_+ t} \end{eqnarray}
Here we've put the $\omega_0$'s back in because we're using $t$ instead of $\xi$.

In the case that $Q \gg 1$ the eigenvalues simplify to \begin{equation}
\lambda_{\pm} = \pm i -\frac{\epsilon}{2} \end{equation}
so that we get \begin{eqnarray}
a^- = \frac{1}{\sqrt{2}}\left[ X + \left( i + \frac{\epsilon}{2} \right) Y \right] &\qquad&
a^+ = \frac{1}{\sqrt{2}}\left[ X + \left(-i + \frac{\epsilon}{2} \right) Y \right] \\
a^-(t) = a^-(0) \exp \left[ \omega_0 \left(-i - \frac{\epsilon}{2}\right) t \right]
& \qquad &
a^+(t) = a^+(0) \exp \left[ \omega_0 \left( i - \frac{\epsilon}{2}\right) t \right] \end{eqnarray}
Note that the time evolution has the same sense from what is typically used in quantum mechanics for the $a$ and $a^{\dagger}$ operators.

\section{Coupled oscillators}

\input{coupled_oscillators.tex}

\subsection{Hamiltonian approach}

\input{coupled_oscillators_hamiltonian.tex}

\section{Coupled transmission lines}

\input{coupled_lines.tex}

\end{document}
