\levelstay{Impulse Response}

The equation for a damped harmonic oscillator is
\begin{equation}
  \ddot{\phi}(t) + \kappa \dot{\phi}(t) + \omega_0^2 \phi(t) = J(t)
\end{equation}
where $\kappa$ has dimensions of $1 / [\text{Time}]$ and the forcing term $J(t)$ has dimensions of $[\phi]/[\text{Time}]^2$.
In the usual way we turn this into an abstract equation
\begin{equation}
  \left( D_t^2 + \kappa D_t + \omega_0^2 \right) \ket{\phi} = \ket{J}
\end{equation}
Our Fourier convention will be
\begin{equation}
  \braket{t}{\omega} = e^{i\omega t}
\end{equation}
With this convention $-iD_t$ is Hermitian.
Going to the $\ket{\omega}$ basis we get
\begin{align}
  \bbraket{\omega}{D_t^2 + \kappa D_t + \omega_0^2}{\phi} &= \braket{\omega}{J} \nonumber \\
  \left( -\omega^2 + i \kappa \omega + \omega_0^2 \right) \ft \phi(\omega) & = \ft J(\omega) \nonumber \\
  \ft \phi(\omega) & = \frac{\ft J(\omega)}{-\omega^2 + i \kappa \omega + \omega_0^2}
  \, .
\end{align}
We study the case where the drive is a single delta function at time $t_0$.
Taking $J(t) = A \delta(t - t_0)$ we get $\ft J(\omega) = A e^{-i \omega t_0}$.
Observe that $A$ has dimensions of $[J] \times [\text{Time}]$.
The denominator factors into $-(\omega - \omega_+)(\omega - \omega_-)$ where
\begin{equation}
  \omega_+ = i \frac{\kappa}{2} + \omega_0' \qquad \omega_{-} = i \frac{\kappa}{2} - \omega_0' \label{eq:omega_plus_minus}
\end{equation}
and
\begin{equation}
  \omega_0' = \omega_0 \sqrt{ 1 - \left( \kappa / 2 \omega_0 \right)^2 } \, .
\end{equation}
Therefore we have
\begin{equation}
  \ft \phi(\omega) = \frac{-A e^{-i \omega t_0}}{(\omega - \omega_{+})(\omega-\omega_{-})} \, .
\end{equation}
In the underdamped case where $\kappa / (2 \omega_0) < 1$, $\omega_0'$ is real.
Upon Fourier transform
\begin{equation}
  \phi(t) = 
  \int_{- \infty}^\infty \frac{d\omega}{2\pi} \ft \phi(\omega) e^{i \omega t}
  = \int \frac{d \omega}{2 \pi} \frac{-Ae^{i \omega(t - t_0)}}{(\omega-\omega_{+})(\omega-\omega_{-})} \,
  \, .
\end{equation}
This integral can be evaluated via contour integration.
Consider first the case $t < t_0$.
In that case we close the contour in the lower half plane.
Since both poles $\omega_\pm$ are in the upper half plane, the integral is zero.
This reflects causality: the oscillator can't know about the impulse before it happens.
For the case $t > t_{0}$ we close in the upper half plane, the contour encloses the poles, and we will get a nonzero result.
Let $\tau \equiv t - t_0$.
Then the residue theorem gives
\begin{align}
  \phi(t) &=
  -A \frac{(2\pi) i }{2 \pi}
  \left(
  \frac{e^{i \omega_+ \tau}}{\omega_+ - \omega_-} +
  \frac{e^{i \omega_- \tau}}{\omega_- - \omega_+}
  \right) \\
  &= \frac{A}{\omega_0'} e^{- \kappa \tau / 2} \sin(\omega_0' \tau)
\end{align}
We see that the system oscillates with frequency $\omega_0'$ and decays with a rate given by the amplitude damping parameter $\kappa$.

\leveldown{Energy decay in the underdamped limit}

In the limit $\kappa \ll \omega_0$ the energy in the system obeys $E(t) = E_0 e^{- \kappa (t - t_0)}$, where $E_0$ is the energy in the system immediately following the drive at time $t_0$.\footnote{See the exercises.}
We define the \newword{quality factor} $Q$ for an oscillating system by
\begin{equation}
  Q \equiv \frac{\textrm{Energy stored}}{\textrm{Energy dissipated per radian}} \, .
\end{equation}
With this definition, we find
\begin{equation}
  Q
  = \frac{E(t)}{-dE/d\,\textrm{radian}}
  = \frac{E(t)}{-(dE/dt)(dt/d\,\textrm{radian})}
  = \frac{E(t)}{\kappa E(t) / \omega_0'}
  \approx \frac{\omega_0}{\kappa}
\end{equation}
where we used the fact that, when $\kappa \ll \omega_0$, $\omega_0' \approx \omega_0$.

